---
title: "Graduate Admissions Prediction"
author: "Kyle Choi"
date: "3/12/2021"
output: html_document
---
***All files related to this project can be found at https://github.com/kc-dev2/grad_admissions_R***

## Purpose

The purpose of this project is to practice exploring and creating models on sample data. More specifically, I have selected graduate admissions data listed on [Kaggle](https://www.kaggle.com/mohansacharya/graduate-admissions). 

While exploring the data, I will ask questions such as "what predictor seems to be most effective in predicting admission probability" and also plot graphs to help communicate/visualize these results. Afterwards, I will create a few linear models to predict admission probabilities and validate these models.

### Exploration

To begin, let us import the data along with necessary libraries:

```{r message=FALSE, results='hide'}
library(tidyverse)

adm <- read_csv("../data/admissions.csv")
```

For data "cleaning", I will check to ensure that there are no missing data. Additionally, any columns that need be renamed or have types changed will be fixed.
```{r adm}
sapply(adm, function(x) sum(is.na(x))) #counting total missing values in each column
str(adm); summary(adm) #checking basic info on data
adm <- rename(adm, GRE = "GRE Score", TOEFL = "TOEFL Score", U_Ranking = "University Rating", adm_chance = "Chance of Admit") #renaming columns
names(adm)
adm$Research <- as.factor(adm$Research) #changing 'Research' column from numerical to factor
str(adm$Research)
```
afdfd this is returning error
```{r adm}
plot(adm) #check relationships between predictors and admission chance percentage
sub_adm <- adm[, !names(adm) %in% c("Research")] #remove "Research" column to allow for correlation matrix
corrplot::corrplot(cor(sub_adm), type = "lower", addCoef.col = "black")
```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
