---
title: "Graduate Admissions Prediction"
author: "Kyle Choi"
date: "3/12/2021"
output: html_document
---
## Purpose

The purpose of this project is to practice exploring and creating models on sample data. More specifically, I have selected graduate admissions data listed on [Kaggle](https://www.kaggle.com/mohansacharya/graduate-admissions). 

While exploring the data, I will ask questions such as "what predictor seems to be most effective in predicting admission probability" and also plot graphs to help communicate/visualize these results. Afterwards, I will create a few linear models to predict admission probabilities and validate these models.

### Exploration

To begin, let us import the data along with necessary libraries:

```{r message=FALSE, results='hide'}
library(tidyverse)

adm <- read_csv("../data/admissions.csv")
```

```{r adm}
summary(adm)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
